Modeling the impacts of replacing car trips with combined public transport and cycling: Reproducible methods, results and actionable evidence from biclaR

### 2023-11-29 paper/PaperCEUS/PaperCEUS.Rmd
xxx and minor improvs
cover letter and title page

---
title: "Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from the biclaR project"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to public transport were also estimated and monetized.
  The results show that 10% of trips could be made by cycling in combination with public transport. Shifting to cycling for the shorter first and last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the public transport leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

@@

To achieve the cycling targets set by the Portuguese national cycling strategy for 2025 and 2030 (4% and 10%, respectively) [@ENMAC], the Lisbon's Metropolitan Department of Transport commissioned *biclaR*[^1], a decision support tool that facilitates the planning, design, and development of a metropolitan cycling network (ref X).
<!-- [@felix2023]. -->

[^1]: See xxx.xxx. *For double blind peer-review process, the authors decided to truncate the name and references to this tool*.
<!-- [^1]: See [biclar.tmlmobilidade.pt](https://biclar.tmlmobilidade.pt/). -->

@@

The jittering pre-processing stage generates a more realistic representation of the trips undertaken than the traditional centroid-based approach but does precisely capture the exactly spatial distribution of trips.
Even where such datasets, they cannot be shared for research due to data privacy regulations.

@@

## Acknowledgements {.unnumbered}
[*blind*]
<!-- This research was funded by the Lisbon's Metropolitan Department of Transport (TML - Transportes Metropolitanos de Lisboa, E.M.T., S.A.), under the *biclaR* Project. -->
<!-- This work is part of the research activity carried out at Civil Engineering Research and Innovation for Sustainability (CERIS) and has been funded by Fundação para a Ciência e a Tecnologia (FCT), Portugal in the framework of project UIDB/04625/2020.  -->
<!-- The authors thank Thomas Götschi (HEAT for Cycling) for providing access to HaaS tool, which is under development. -->

>>>>

---
title: "Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from biclaR"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to public transport were also estimated and monetized.
  The results show that 10% of trips could be made by cycling in combination with public transport. Shifting to cycling for the shorter first and last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the public transport leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

@@

To achieve the cycling targets set by the Portuguese national cycling strategy for 2025 and 2030 (4% and 10%, respectively) [@ENMAC], the Lisbon's Metropolitan Department of Transport commissioned *biclaR*[^1], a decision support tool that facilitates the planning, design, and development of a metropolitan cycling network [@felix2023].

<!-- [^1]: See xxx.xxx. *For double blind peer-review process, the authors decided to truncate the name and references to this tool*. -->
[^1]: See [biclar.tmlmobilidade.pt](https://biclar.tmlmobilidade.pt/).

@@

The jittering pre-processing stage generates a more realistic representation of the trips undertaken than the traditional centroid-based approach but does precisely capture the exactly spatial distribution of trips.
Even where such datasets exist, they cannot be shared for research due to data privacy regulations.

@@

## Acknowledgements {.unnumbered}
[*blind*]
<!-- This research was funded by the Lisbon's Metropolitan Department of Transport (TML - Transportes Metropolitanos de Lisboa, E.M.T., S.A.), under the *biclaR* Project. -->
<!-- This work is part of the research activity carried out at Civil Engineering Research and Innovation for Sustainability (CERIS) and has been funded by Fundação para a Ciência e a Tecnologia (FCT), Portugal in the framework of project UIDB/04625/2020. -->
<!-- The authors thank Thomas Götschi (HEAT for Cycling) for providing access to _HaaS_ tool, which is under development. -->

### 2023-11-29 paper/PaperCEUS/PaperCEUS.tex
xxx and minor improvs
cover letter and title page

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Modeling the impacts of replacing car trips with combined public transport and cycling: Reproducible methods, results and actionable evidence from the XXX project},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

@@

  \title{Modeling the impacts of replacing car trips with combined
public transport and cycling: Reproducible methods, results and
actionable evidence from the XXX project}
    \author[CERIS]{Rosa Félix\corref{cor1}%
  \corref{cor1}%
  }
   \ead{rosamfelix@tecnico.ulisboa.pt} 
    \author[CERIS]{Filipe Moura%
  %
  }
   \ead{fmoura@tecnico.ulisboa.pt} 
    \author[ITS]{Robin Lovelace%
  %
  }
   \ead{r.lovelace@uleeds.uk} 
      \affiliation[CERIS]{CERIS, Instituto Superior Técnico -
Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal}
    \affiliation[ITS]{Institute for Transport Studies, University of
Leeds. 34-40 University Rd, Leeds LS2 9JT, UK}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  A high proportion of car trips can be replaced by a combination of
  public transit and cycling for the first-and-last mile. This paper
  estimates the potential for cycling combined with public transit (PT)
  as a substitute for car trips in the Lisbon metropolitan area and
  assesses its socio-environmental impacts using open data and open
  source tools. A decision support tool that facilitates the design and
  development of a metropolitan cycling network was developed
  (\emph{XXX}). The social and environmental impacts were assessed using
  the \emph{HEAT for Cycling} and the \emph{HEAT as a Service} tools.
  The impacts of shifting car trips to PT were also estimated and
  monetized. The results indicate that 10\% of trips could be made by
  bicycle + PT combination. Shifting to cycling for the first-and-last
  mile stages can reduce annual CO\textsubscript{2}eq emissions from
  3,000 to 7,500 tons/day, while for the PT leg, the transfer from car
  avoids of up to 20,500 tons of CO\textsubscript{2}eq emissions per
  year. The estimated socio-environmental benefits are of €125 million
  to €325 million over 10 years. This evidence can support policymakers
  to prioritize interventions that reduce the reliance on private motor
  vehicles.\\
  \end{abstract}
    \begin{keyword}
    Active transport \sep Intermodality \sep First and last
mile \sep Health economic assessment \sep Environmental impacts \sep 
    Open data and methods
  \end{keyword}

@@

Combining public transportation (PT) and cycling for the first and last
mile in metropolitan areas can significantly replace private car trips
(Martens, 2007; Rietveld, 2000). This approach requires interventions
and programs to make bicycling more appealing, and the resulting public
investments can have significant social and environmental benefits.

According to the latest mobility survey conducted in 2018 (INE, 2018),
the LMA registered a total of 5.3 million daily trips, with only 0.5\%
by bicycle. Car modal share was 58.4\%, while PT accounted for 15.5\%
(see Figure \ref{fig:mododist}). The number of intra-municipal trips ---
with origin and destination in the same municipality --- amounts to 3.5
million trips. This exceeds the number of inter-municipal trips (1.8
million trips), involving travel between different municipalities. Cars
and public transport are the most used modes for intercity trips, with
cars being the predominant choice for all journeys.\\
53\% trips are up to 5 km distance, and 71\% up to 10 km. Nevertheless,
29\% of trips are longer than 10 km, which requires the use of motorized
modes, or active modes in combination with public transport.

@@

To achieve the cycling targets set by the Portuguese national cycling
strategy for 2025 and 2030 (4\% and 10\%, respectively) (Presidência do
Conselho de Ministros, 2019), the Lisbon's Metropolitan Department of
Transport introduced \emph{XXX}\footnote{See xxx.xxx. \emph{For double
  blind peer-review process, the authors decided to truncate the name
  and references to this tool}.}, a decision support tool that
facilitates the planning, design, and development of a metropolitan
cycling network (ref X).

\emph{XXX} builds on the Propensity to Cycle Tool\footnote{See
  \href{https://www.pct.bike/}{pct.bike}.} (PCT), a web application and
research project funded by the UK's Department for Transport in 2015
which launched nationally in 2017 as part of the government's Cycling
and Walking Investment Strategy. The PCT initially used only
origin-destination data for commuting trips as the basis of estimates of
cycling potential at zone, route and route network levels (Lovelace et
al., 2017). The PCT has been extended to include cycling potential for
travel to school in England (Goodman et al., 2019) and other trip types
in other countries.\footnote{See \href{https://www.npt.scot}{npt.scot}
  and \href{https://cruse.bike}{cruse.bike} for examples of the PCT in
  Scotland and Ireland that include estimates of cycling for other
  purposes.} However, to the best of our knowledge, this is the first
time that the method has been integrated with public transport data
using multi-modal routing to estimate the potential and benefits of
multi-stage cycling and PT trips.

@@

The mobility survey data (INE, 2018) is the basis for this project and
defines the baseline scenario. Despite being conducted in the
pre-pandemic period (2017), this dataset represents the most
comprehensive and up-to-date information on urban mobility in Portuguese
metropolitan areas (Lisbon and Porto).

We used a method for disaggregating the origins and destinations of
trips between the centroids of two districts (same as ``parish'') to
ensure that a district is not solely characterized by a single point of
origin and destination for its trips. Aggregating all trips into
centroids renders the exercise less realistic, as it excludes a
significant portion of short-distance trips, a prevalent characteristic
of active travel (Lovelace et al., 2022b). The OD Jittering method
breaks down a single point (i.e., the centroid of an area) into multiple
random points on the existing and neighboring road network, using
OpenStreetMap as a reference. This method then distributes the volume of
trips within the district among the randomly generated
origin-destination pairs.

Using the
\href{https://github.com/dabreegster/odjitter}{\texttt{odjitter} R
package}, we employed a maximum disaggregation level of 100 trips per
O-D pair for this project. Figure \ref{fig:jitter} illustrates the
contrast between trip representation through the traditional method,
which connects a single desire line between each district, and the
presentation achieved through the randomization and disaggregation of
trips between districts, specifically for the Lisbon metropolitan area.

@@

Although this method provides a more realistic representation of the
trips undertaken compared to the traditional approach, it does not fully
align with the actual O-D pairs of trips, which remain unknown due to
data privacy regulations.

@@

The intermodality scenario considers trips combining PT and cycling for
the first and last legs. In a conservative approach, we have restricted
our analysis to the first and last legs with a combined length of up to
5 km (for example: 1 km from origin to interface \emph{A} plus 4 km from
interface \emph{B} to destination) or up to 25 minutes on bike
travel-time. Furthermore, we have imposed restrictions on PT usage,
limiting it to trips without PT transfers, and within a duration of up
to 2 hours (120 minutes). Additionally, we have only included PT modes
that can easily accommodate bicycles, such as trains, ferries, trams,
and inter-municipal bus lines equipped with bike racks (Figure
\ref{fig:map1}). These restrictions can be eased in the future when
testing more developed policy interventions to enhance intermodality
between cycling and PT, considering both the vehicle and infrastructure
perspectives.

@@

Table \ref{tab:summary22} presents the results of the avoided emissions
and its monetization for the second leg of the journey, by replacing car
trips with potential TP trips.\\
Regarding the PT segment, the shift from private car would lead to the
mitigation of CO\textsubscript{2} equivalent emissions to 8,500 to
20,800 tons annually, valued in €1.4 million to €3.5 million yearly.

@@

Shifting from car to cycling + PT can reduce annual
CO\textsubscript{2}eq emissions by 14,000 to 36,000 tons per year. The
10-year socio-environmental benefits account for €125 million to €325
million, depending on the cycling targets.

@@

Our findings suggest that cycling \emph{in combination} with PT could
potentially replace 10\% of current LMA trips, with an additional 6\% of
PT journeys prone to further substitution.

@@

It has become progressively more common to establish strategic plans, at
national, regional or municipal level, to mitigate climate change. Among
these, the Sustainable Urban Mobility Plans (SUMP\footnote{See
  \href{https://www.eltis.org/mobility-plans/sump-concept}{eltis.org/mobility-plans/sump-concept}}),
promoted by the European Commission, are becoming popular in Europe,
although authorities are designing documents of this sort all over the
world. The definition of targets associated with a timeframe for
reducing dependence on the individual motorized vehicle, or targets for
the use of active modes such as walking and cycling, are too often not
accompanied by estimates of their social, environmental and economic
impacts. It is important for authorities and practitioners to know how
to estimate those impacts, which tools are available to support them in
the process, and what results to expect.

This paper attempts to estimate the socio-environmental impacts
associated with the modal shift from car to intermodal trips by bicycle
in combination with public transport, in the Lisbon metropolitan area,
in order to meet the national targets set for bicycle use of 4\% and
10\% by 2025 and 2030, respectively.

The information on socio-economic benefits can support policy-makers in
prioritizing interventions to reduce the reliance on individual
motorized transportation, and to better communicate their decisions by
providing the expected avoided GHG and air pollutant emissions, and the
monetized socio-economic benefits for short and long terms.

The information available at \emph{XXX} tool -- an open access website
-- can be downloaded and used with any GIS software. This allows
practitioners to, for example, gain insights into which potential
cycling connections have the highest socio-environmental impacts,
quantified in tons of avoided CO\textsubscript{2}eq emissions, or in
long term social benefits.

We used open-source tools and transparent methods. By making the
research process publicly accessible in a code repository (see
\protect\hyperlink{research-data}{Research Data}), it enables the
replication of similar estimates for socio-environmental impacts,
resulting from a modal shift from car to bicycle in combination with PT,
in other metropolitan areas.

@@

The data and the code to produce XXX and this paper are available at:
https://github.com/{[}\emph{blind}{]}.

@@

@@

@@

@@

>>>>

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Modeling the impacts of replacing car trips with combined public transport and cycling: Reproducible methods, results and actionable evidence from biclaR},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

@@

  \title{Modeling the impacts of replacing car trips with combined
public transport and cycling: Reproducible methods, results and
actionable evidence from biclaR}
    \author[CERIS]{Rosa Félix\corref{cor1}%
  \corref{cor1}%
  }
   \ead{rosamfelix@tecnico.ulisboa.pt} 
    \author[CERIS]{Filipe Moura%
  %
  }
   \ead{fmoura@tecnico.ulisboa.pt} 
    \author[ITS]{Robin Lovelace%
  %
  }
   \ead{r.lovelace@uleeds.uk} 
      \affiliation[CERIS]{CERIS, Instituto Superior Técnico -
Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal}
    \affiliation[ITS]{Institute for Transport Studies, University of
Leeds. 34-40 University Rd, Leeds LS2 9JT, UK}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  A high proportion of car trips can be replaced by a combination of
  public transit and cycling for the first-and-last mile. This paper
  estimates the potential for cycling combined with public transit as a
  substitute for car trips in the Lisbon metropolitan area and assesses
  its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of
  a metropolitan cycling network was developed (\emph{biclaR}). The
  social and environmental impacts were assessed using the \emph{HEAT
  for Cycling} and the \emph{HEAT as a Service} tools. The impacts of
  shifting car trips to public transport were also estimated and
  monetized. The results show that 10\% of trips could be made by
  cycling in combination with public transport. Shifting to cycling for
  the shorter first and last mile stages can reduce annual
  CO\textsubscript{2}eq emissions from 3,000 to 7,500 tons/day, while
  for the public transport leg, the transfer from car avoids of up to
  20,500 tons of CO\textsubscript{2}eq emissions per year. The estimated
  socio-environmental benefits are of €125 million to €325 million over
  10 years. This evidence can support policymakers to prioritize
  interventions that reduce the reliance on private motor vehicles.\\
  \end{abstract}
    \begin{keyword}
    Active transport \sep Intermodality \sep First and last
mile \sep Health economic assessment \sep Environmental impacts \sep 
    Open data and methods
  \end{keyword}

@@

Combining public transportation (PT) and cycling for the first and last
mile in metropolitan areas can replace a high proportion of private car
trips (Martens, 2007). In The Netherlands, which has the highest mode
share of cycling in the world, cycling accounts for more than a third of
all trips to and from rail stations at the `home' end of the journey,
greatly increasing the ability of the transport system (Rietveld, 2000).
This approach to reducing car dependency and associated externalities
requires interventions and programs to make bicycling more appealing (La
Paix et al., 2021). The resulting public investments can have
significant social and environmental benefits (International Transport
Forum, 2017). Despite the benefits of cycling-PT intermodality, the
potential of this combination is often overlooked in transport planning
(La Paix et al., 2021).

The potential of cycling as a complementary mode of PT is substantial
worldwide, especially in cities with established public transport
networks or substantial ambitions to develop them. In the Lisbon
metropolitan area (LMA) the largest metropolitan area in Portugal, the
modal share of cycling is low, but the potential for cycling as a
complementary mode of PT is high. According to the latest mobility
survey conducted in 2018 (INE, 2018), the LMA registered a total of 5.3
million daily trips, with only 0.5\% by bicycle. Car modal share was
58.4\%, while PT accounted for 15.5\% (see Figure \ref{fig:mododist}).
The number of intra-municipal trips --- with origin and destination in
the same municipality --- amounts to 3.5 million trips. This exceeds the
number of inter-municipal trips (1.8 million trips), involving travel
between different municipalities. Cars and public transport are the most
used modes for intercity trips, with cars being the predominant choice
for all journeys.\\
53\% trips are up to 5 km distance, and 71\% up to 10 km. Nevertheless,
29\% of trips are longer than 10 km, which requires the use of motorized
modes, or active modes in combination with public transport.

@@

To achieve the cycling targets set by the Portuguese national cycling
strategy for 2025 and 2030 (4\% and 10\%, respectively) (Presidência do
Conselho de Ministros, 2019), the Lisbon's Metropolitan Department of
Transport commissioned \emph{biclaR}\footnote{See
  \href{https://biclar.tmlmobilidade.pt/}{biclar.tmlmobilidade.pt}.}, a
decision support tool that facilitates the planning, design, and
development of a metropolitan cycling network (Félix et al., 2022).

\emph{biclaR} builds on the Propensity to Cycle Tool\footnote{See
  \href{https://www.pct.bike/}{pct.bike}.} (PCT), a web application and
research project funded by the UK's Department for Transport in 2015
which launched nationally in 2017 as part of the government's Cycling
and Walking Investment Strategy. The PCT initially used only
origin-destination data for commuting trips as the basis of estimates of
cycling potential at zone, route and route network levels (Lovelace et
al., 2017). The PCT has been extended to include cycling potential for
travel to school in England (Goodman et al., 2019) and other trip types
in other countries.\footnote{See \href{https://www.npt.scot}{npt.scot}
  and \href{https://cruse.bike}{cruse.bike} for examples of the PCT in
  Scotland and Ireland that include estimates of cycling for other
  purposes.} However, to the best of our knowledge, this is the first
time that the method has been integrated with public transport data
using multi-modal routing to estimate the potential and benefits of
multi-stage cycling and PT trips.

@@

The mobility survey data (INE, 2018) is the basis of the baseline
scenario and trip rates presented in this paper. Conducted in the
pre-pandemic period (2017), this OD dataset represents the most
comprehensive and up-to-date information on urban mobility in Portuguese
metropolitan areas (Lisbon and Porto).

We used `jittering' to disaggregate the OD data, resulting in a wide
spatial distribution trip origins and destinations (Lovelace et al.,
2022b). The method works by sampling `sub-points' (nodes on the
transport network represented in OpenStreetMap in this case) and using
these instead of a single point (typically the centroid) to represent
trip origins and destinations for each zone. This method then
distributes the trips to desire lines connecting the subpoints based on
a `disaggregation threshold' which determines the maximum number of
trips that can be represented by a single desire line.

Using the
\href{https://github.com/dabreegster/odjitter}{\texttt{odjitter} R
package}, we disaggregated the OD data into desire lines reprenting a
maximum of 100 trips each. Figure \ref{fig:jitter} illustrates the
contrast between trip representation through the traditional method,
which connects a single desire line between each district, and the
presentation achieved through the randomization and disaggregation of
trips between districts, specifically for the Lisbon metropolitan area.

@@

The jittering pre-processing stage generates a more realistic
representation of the trips undertaken than the traditional
centroid-based approach but does precisely capture the exactly spatial
distribution of trips. Even where such datasets exist, they cannot be
shared for research due to data privacy regulations.

@@

The intermodality scenario considers trips made by PT in which cycling
is used for the first and last legs. We restricted our analysis to the
first and last legs with a combined length of up to 5 km (for example: 1
km from origin to interface \emph{A} plus 4 km from interface \emph{B}
to destination) or up to 25 minutes on bike travel-time. This
conservative approach was adopted to capture the fact that cycling
stages as part of a multi-modal trip are likely to be shorter than
cycling-only trips (van Mil et al., 2021). Furthermore, we have imposed
restrictions on PT usage, limiting it to trips without PT transfers, and
within a duration of up to 2 hours (120 minutes). Additionally, we have
only included PT modes that can easily accommodate bicycles, such as
trains, ferries, trams, and inter-municipal bus lines equipped with bike
racks (Figure \ref{fig:map1}). These restrictions can be eased in the
future when testing more developed policy interventions to enhance
intermodality between cycling and PT, considering both the vehicle and
infrastructure perspectives.

@@

Table \ref{tab:summary22} presents emissions reductions and associated
economic benefits associated with the second (PT) leg of trips. The
shift from private car associated with thes PT segments would reduce
CO\textsubscript{2} equivalent emissions by 8,500 to 20,800 tons
annually, valued in €1.4 million to €3.5 million yearly, for the 4\% and
10\% targets, respectively.

@@

Shifting from car to cycling in combination with PT can reduce annual
CO\textsubscript{2}eq emissions by 14,000 to 36,000 tons per year. The
10-year socio-environmental benefits account for €125 million to €325
million, depending on the cycling targets.

@@

Our findings show that cycling \emph{in combination} with PT could
replace 10\% of current LMA trips, with an additional 6\% of PT journeys
prone to further substitution, based on conservative assumptions.

@@

This paper estimated the potential for combining cycling and PT to
substitute car trips in the LMA, while achieving the national cycling
targets and supporting decarbonization goals. It has become
progressively more common to establish strategic plans, at national,
regional or municipal level, to mitigate climate change. Among these,
the Sustainable Urban Mobility Plans (SUMP\footnote{See
  \href{https://www.eltis.org/mobility-plans/sump-concept}{eltis.org/mobility-plans/sump-concept}}),
promoted by the European Commission, are becoming popular in Europe,
although authorities are designing documents of this sort all over the
world. The definition of targets associated with a timeframe for
reducing dependence on the individual motorized vehicle, or targets for
the use of active modes such as walking and cycling, are too often not
accompanied by estimates of their social, environmental and economic
impacts. It is important for authorities and practitioners to know how
to estimate those impacts, which tools are available to support them in
the process, and what results to expect.

This paper quantifies the benefits of replacing car trips with cycling,
in combination with public transport. The case study of the Lisbon
metropolitan area demonstrates that cycling-PT integration can help meet
the national targets set for bicycle use of 4\% and 10\% by 2025 and
2030, respectively.

The quantification of benefits can support policy-makers in prioritizing
interventions to reduce the reliance on individual motorized
transportation. The presentation of the results in an open access web
application will help communicate decisions. Furthermore, the provision
of datasets resulting from this project provides a foundation for
further research and development of new tools and methods. The methods
are reproducible and based on open source software, which can be applied
to other cities and metropolitan areas, supporting the decarbonization
of transport systems internationally.

@@

The data and the code to reproduce the results will be made available
upon publication.

@@

\leavevmode\vadjust pre{\hypertarget{ref-felix2023}{}}%
Félix, R., Lovelace, R., \& Moura, F. (2022). \emph{{biclaR - Ferramenta
de apoio ao planeamento da rede ciclável na área metropolitana de
Lisboa}}. {CERIS - Instituto Superior Técnico and Transportes
Metropolitanos de Lisboa}. \url{https://biclar.tmlmobilidade.pt}

@@

\leavevmode\vadjust pre{\hypertarget{ref-internationaltransportforum_integrating_2017}{}}%
International Transport Forum. (2017). \emph{Integrating {Urban Public
Transport Systems} and {Cycling}}. {Organisation for Economic
Co-operation and Development}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-lapaix_role_2021}{}}%
La Paix, L., Cherchi, E., \& Geurs, K. (2021). Role of perception of
bicycle infrastructure on the choice of the bicycle as a train feeder
mode. \emph{International Journal of Sustainable Transportation},
\emph{15}(6), 486--499.
\url{https://doi.org/10.1080/15568318.2020.1765223}

@@

\leavevmode\vadjust pre{\hypertarget{ref-vanmil_insights_2021}{}}%
van Mil, J. F. P., Leferink, T. S., Annema, J. A., \& van Oort, N.
(2021). Insights into factors affecting the combined bicycle-transit
mode. \emph{Public Transport}, \emph{13}(3), 649--673.
\url{https://doi.org/10.1007/s12469-020-00240-2}
